Advanced Normalization Tools (ANTS)

### 2009-07-08 Documentation/Tex/ants.tex
Changed the figure labels to appear after the captions which seems to have fixed the reference problem.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@47 fd2da0a4-d968-48df-9cb5-d8c496287cbf
ok
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@53 fd2da0a4-d968-48df-9cb5-d8c496287cbf

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\label{fig:cmd}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.  The 
SyN option implements the method in \cite{Avants2008b} and evaluated in 
\cite{Klein2009}.}
\vspace{-0.1in}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  You
may ``check'' for consistency between two image header definitions by
using: ImageMath 3 Image2repaired.nii.gz CompareHeadersAndImages
Image1.nii.gz Image2.nii.gz \newline This will compare Image2 to
Image1 and ``fix'' the header of Image2, writing out to
Image2repaired.

@@

\subsubsection{Linear Registration}
\begin{figure}
\label{fig:aff}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSanatomy.pdf} 
\itkcaption[ANTS Introduction]{The anatomy of an ANTS optimization and application of the resulting warping.}
\vspace{-0.1in}
\end{figure}
The most basic mapping between images is an affine mapping. The ANTS 
affine mapping syntax is shown in the affine mapping figure~\ref{fig:aff}. The
parameter breakdown of the ANTS call in the affine figure is:
\begin{itemize}
\item  ANTS -- the normalization program
\item 2 -- the expected image dimension
\item -m MI[...,...,1,32] -- the similarity metric "mutual information" with a 32-bin square joint histogram and weight of 1.  
Metric details will be discussed in a later section. 
\item -i 0 -- no deformable iterations (affine only).
\item -o ab -- the output prefix.  A specific extension may also be specified, e.g.   -o ab.mhd  .
\end{itemize}
The parameter breakdown of the WarpImageMultiTransform call in the affine figure is:
\begin{itemize}
\item WarpImageMultiTransform -- to apply the mapping output from ANTS to an image.
\item 2 -- the expected image dimension.
\item the image to be deformed is input next.
\item the deformed image output file name is next.
\item the transform itself is next -- if is affine and preceded by " -i ", we apply the inverse affine map.
\item the " -R " option dictates the domain you want to warp into -- usually the ``fixed'' image, unless using the inverse mapping, in which case one switches the role of fixed and moving images. 
\end{itemize}
Note that -i 0 (in the ANTS call) means that no deformable mapping is performed. The call "ANTS 2" means to treat the images as 2-dimensional. "ANTS 3" would be called for 3D images. The " -m " term defines the similarity metric. 
Note that no more than one image interpolation is used during optimization.  
That is, ANTS always references the original image before applying a transformation or series of transformations.

@@

\subsubsection{Deformable Registration}
Affine mapping is adequate when the
difference between images involves only rotation, scaling or shearing.
Other data requires more deformable mappings to capture shape
differences and find a good alignment of image features.
Figure~\ref{fig:antshier} shows how deformable mapping may improve the
correspondence of the deformed beetle to the ford image.
\begin{figure}
\label{fig:antshier}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSElastToDiff.pdf} 
\itkcaption[ANTS Transformation Examples]{This example shows the degree to which the beetle (b.img) may be deformed to the ford (a.img) under different transformation models.  Left to right increases the degrees of freedom in the mapping and thus the registration accuracy.}
\vspace{-0.1in}
\end{figure}
Most ANTS-based applications use symmetric diffeomorphic normalization.  However, 
ANTS also enables a simpler parameterization of a deformable mapping via a regularized 
vector space.  We term these types of transformations as ``elastic''.  The original Demons algorithm 
provides a classic example of using a regularized vector space for nonlinear registration.  
Caveats are that a regularized vector space may not preserve the underlying topology 
and may also prove too inflexible to capture the shape changes one is after.  Both of these 
shortcomings motivate the use of diffeomorphisms.

\noindent{\bf Elastic/Vector Space Transformations.} If we assume no affine transformation, then 
an elastic transformation involves computing a mapping from image $I(\x)$ to image $J(\x)$ 
through a deformation field $\disp( \x )$.  
\begin{figure}
\label{fig:pull}
\center \includegraphics[width=0.75\textwidth]{./Figures/PullBack.pdf} 
\itkcaption[ANTS Deformation]{A deformation in a digital domain.}
\end{figure}  
The deformation is defined in the physical space of the image 
and dictates the positional difference between corresponding features in the two images.  
Thus, if a feature defined at $I(\x)$ matches a feature in $J$ at position $\y$ then the deformation 
field at $\x$ should give $\disp ( \x )=\y-\x$.   
Such a deformation field may be applied to deform 
image $J$ into image $I$ by composing the mapping $J_{deformed}(\x)=J(\x+\disp(\x))$.  
In a perfect world, then $I(\x)=J_{deformed}(\x)$, though this is rarely the case.  Figure~\ref{fig:pull} 
visualizes the deformation of an image under this standard model. 
Gradient descent optimization of an elastic mapping may be summarized (crudely) as: 
\begin{align}
\text{Compute the similarity gradient:   }  \nabla E = \partial_{\disp} \Pi(I,J(\x+\disp(\x))). \nonumber \\
\text{Update the deformation field:  }  \disp(\x) \leftarrow \disp(\x)+\delta \nabla E. \nonumber \\ 
\text{Regularize the deformation field:  }  \disp(\x) \leftarrow G_\sigma \star \disp(\x),
\end{align}
where $\Pi$ is the similarity, $\delta$ is a gradient step length and $G_\sigma$ is a gaussian smoother.
This optimization is captured in the following ANTS command:
\begin{verbatim}
ANTS 3 -m PR[AtlasHead.nii,crAnatomical.nii,1,4] -r Gauss[0,3]  
-o ElasticTest.nii.gz -i 30x20x10  -t Elast[1.5] 
\end{verbatim}
Here, we use the PR metric (a cross-correlation implementation) with
window radius 4, weight 1 and gradient step length 1.5.  The
optimization will be performed over three resolutions with a maximum
of 30 iterations at the coarsest level, 20 at the next coarsest and 10
at the full resolution.  We use a Gaussian regularizer with a sigma of
3 that operates only on the deformation field and not on the similarity gradient, as 
0 is passed as the first parameter to \verb Gauss .  One may see the
correspondence, yet again, between the ANTS call and the optimization
scheme.  The optimization will stop when either the energy cannot be 
smoothly minimized or the maximum number of iterations is reached. 
BSpline regularization is also available in ANTS -- see the DMFFD 
section below.

\noindent{\bf Warping and Invertibility.}  ANTS uses physical space to
define mappings. The origin etc is in the coordinates of the world in
which the picture (e.g. MRI) was taken. Then, warping between physical
coordinates is relatively easy. Differences in bounding boxes etc
present no problem -- assuming you avoid inconsistent headers
i.e. origins, directions, data orientation. One may use PrintHeader to
check the data and run simple, fast tests (few iterations) to perform
sanity checks before running through loads of data.
An ANTS deformation consists of
a standard naming prefix plus a standard naming extension. We usually 
assume nii but other file types may be used. The standard naming
extension is Warpxvec.nii for the x component of the deformation with
similar naming for y and z. Each component is stored in a separate
scalar image. The value of a voxel of a deformation/warp component
image is the physical space displacement from that voxel. Note that
the inverse mapping is stored in InverseWarpxvec.nii and provides the
mapping in the opposite direction.  Note that an inverse -- in the 
digital domain -- is only approximate as shown in figure~\ref{fig:inv}.
\begin{figure}
\label{fig:inv}
\center \includegraphics[width=0.75\textwidth]{Figures/DeformationsInDigitalDomain.pdf} 
\itkcaption[Digital Invertibility]{Digital invertibility presents some limitations. 
Here, we see that invertibility is not exact but is gained only by interpolation.  
Thus, in three-dimensional scenarios in particular, there are fundamental limits to the 
degree of invertibility that may be achieved.  The second and third voxels -- from left -- 
in image A undergo an expansion by a factor of 2.  That is, under the mapping, 
2 voxels are mapped to 4.  This gives the definition of the Jacobian -- computed 
by  ANTSJacobian -- which is a unitless measure defined by the 
ratio of volumes.  Thus, $\mathcal{J}(\x)=V(\phi(\x))/V(\x)$ where $V$ represents 
the volume operation and $\x$, here, may be a small object.  Thus, if $\phi$ -- the mapping -- 
causes expansion, then $\mathcal{J}(\x) > 1$. }
\vspace{-0.1in}
\end{figure}  
A few important notes follow. 
(1) Deformation directionality: Warps/deformations applied to images occur in the opposite direction from warps/deformations applied to points. That is, if we map image B to A using 
\begin{verbatim}
ANTS 2 -m PR[A,B,1,2] -o OUTPUT 
\end{verbatim}
 then we get a deformation called \verb OUTPUTWarp+extensions ~and an affine transform called \verb OUTPUTAffine.txt . 
This composite mapping - when applied to B - will transform B into the space of A. 
However, if we have points defined in B that we want to map to A, 
we have to use \verb OUTPUTInverseWarp ~and the inverse of \verb OUTPUTAffine.txt .  
This is because the domain and range of the map's definition need to be dense in a way that is appropriate for the data type. 
Figure~\ref{fig:inv} illustrates the concept. 
(2) Older Image Formats: older image formats (e.g. Analyze) may not
have proper origin/offset/direction support! In these cases, we
recommend converting to nii and verifying that data overlays properly.
(3) Transorm Composition: Composition of transforms may be achieved with
ComposeMultiTransform.  (4) Warping with inverses and concatenations --
viable when using diffeomorphisms -- are described in
figure~\ref{fig:warp}.
 \begin{figure}
\label{fig:warp}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax2.pdf} 
\itkcaption[ANTS Warping]{Here, we detail how one applies 
a deformation and associated inverses.  We also see how 
WarpImageMultiTransform can be used to concatenate a series 
of transformations.}
\vspace{-0.1in}
\end{figure}

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\label{fig:antssyn}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization \cite{Avants2008b,Kim2008} -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

@@

\subsection{Diffusion Tensor Normalization}
The ImageMath program -- via TensorFA and TensorMeanDiffusion -- 
may derive scalar images from DTI.  Such images may be used to 
distortion correct DTI to T1 or to map DTI together.   See the 
ANTS/Scripts program called \verb antswithdt.sh ~for a 
ready to go example.   Figure~\ref{fig:dterr} shows what might 
happen if your tensor entries are stored in the wrong order. 
ANTS expects the nifti standard ordering of the DT six vector.
\begin{wrapfigure}{r}{0.5\textwidth}
\label{fig:dterr}
\includegraphics[width=0.5\textwidth]{Figures/dtiordererr.pdf} 
\itkcaption[DT Entry Error]{The FA and mean diffusion of the tensor may be corrupted, 
as at right, if the order of the DT entries is wrong.}
\vspace{-0.1in}
\end{wrapfigure}  
\subsection{Multivariate Normalization with ANTS}
\begin{figure}
\label{fig:mvnorm}
\includegraphics[width=0.9\textwidth]{Figures/multivariatenorm.pdf} 
\itkcaption[Multivariate Normalization]{The current version of ANTS 
may perform normalization of multiple modalities by combining {\em intrasubject, intermodality} mappings with {\em intersubject, intramodality} maps to a 
group template.  In brain imaging, the intersubject maps are usually guided by the T1 component.}\vspace{-0.1in}
\end{figure}
Multivariate normalization may be performed in two ways with ANTS.  
First, as shown in figure~\ref{fig:mvnorm}, intra-subject mappings may 
be used to conform all modalities into a common subject space defined by 
the ``anchor'' modality.  In brain imaging, this is usually the T1 component.

@@

The steps are: \begin{figure}

@@

\itkcaption[ANTS Optimal Template]{This example may be recreated by the user.  The shared anatomy -- 

@@

\subsection{User/Label-Guided Normalization for Hippocampus Mapping}
See the link for details on hippocampus labeling:
\href{http://picsl.upenn.edu/ANTS/hipptutorial.php}{http://picsl.upenn.edu/ANTS/hipptutorial.php}.
\noindent{\bf Expectation-Based Point-Set Registration.}
Here, we apply the expectation-based point set registration method 
for mapping labeled points sets, as described in \cite{Pluta2008}.   ITK-SNAP may be used to label 
images and exported segmentation images may be input to the 
PSE metric below, as labeled data.  The Frown and Smile data is used 
as example.  This data is available in the 
ANTS/Examples/Data/ directory.
\begin{figure}
\label{fig:frown}
\includegraphics[width=0.45\textwidth]{Figures/frowntosmile.png}
\includegraphics[width=0.45\textwidth]{Figures/smiletofrown.png}
\itkcaption[Frown To Smile]{Expected output for the frown to smile 
shows a smooth, though large deformation.  The grids are overlaid on 
the deformed images. }
\end{figure}
\begin{verbatim}
ANTS 2 -o PSEtest  -i 91x70x55x40x30  -r Gauss[3,0.] -t SyN[0.2]
   -m  PSE[Frown.nii,Smile.nii,Frown.nii,Smile.nii,0.75,0.1,11,0,10] 
   -m  MSQ[Frown.nii,Smile.nii,1,0]    --number-of-affine-iterations 0

@@

The ANTS toolkit handles this situation by a constrained cost-function
masking approach. In short, the mapping within an excluded region (as
determined by an inclusive mask) is determined by the solution outside
of the region. To achieve this with ANTS, one would use a call similar
to this:
\begin{verbatim}
ANTS 3 -m PR[tp22_s1.nii,template.nii.gz,1,4] -i 50x20x0
   -o tp22map -t SyN[0.5] -x mask.nii.gz -r Gauss[3,0] 
\end{verbatim}
The additional parameter needed for this approach 
is the \verb -x ~ \verb mask.nii.gz ~ which 
is the mask defined in the "fixed" (or individual) image space.  
Mask values above 0.5 are considered non-outlier or non-lesioned regions.
The labeling may be performed with ITK-SNAP. 
Figure~\ref{fig:lesion} illustrates the approach.
\begin{figure}
\label{fig:lesion}
\includegraphics[width=1\textwidth]{Figures/lesionstudy.pdf}
\itkcaption[ANTS Lesions]{Here, we see a semi-automated approach for lesion analysis
based on diffeomorphic normalization.  The user outlines the lesion
site in the subject space, as shown in (a).  Diffeomorphic mapping is
then used to deform the subject image in (a) into the space of the
reference template.  The deformed subject is shown in panel (b).
Panel (c) shows the reference template space with the estimated
position of the subject's lesion overlaid on the healthy tissue of the
template.  This approach enables one to compare the subject image
against prior information stored in the template space such as
anatomical labels or statistics on the appearance of lesioned and/or
normal tissue.}
\end{figure}
Once the ANTS mapping is solved, 
one is able to estimate the location of the lesion in the template space by mapping the inverse lesion mask to the template.
To take the negative image of the inclusive mask 
and warp the mask to the template :
\begin{verbatim}
ImageMath 3 lesionmask.nii.gz Neg inclusivemask.nii.gz
WarpImageMultiTransform 3 lesionmask.nii.gz lesiontotemplate.nii.gz 
   -R template.nii.gz -i tp22mapAffine.txt tp22mapInverseWarp.nii
\end{verbatim}
More details on this approach are being prepared for publication.
See also: \verb http://www.ncrrn.org/papers/methodology_papers/sp_norm_kim.pdf .

>>>>

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.  The 
SyN option implements the method in \cite{Avants2008b} and evaluated in 
\cite{Klein2009}.}
\vspace{-0.1in}
\label{fig:cmd}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  You
may ``check'' for consistency between two image header definitions by
using: ImageMath 3 Image2repaired.nii.gz CompareHeadersAndImages
Image1.nii.gz Image2.nii.gz \newline This will compare Image2 to
Image1 and ``fix'' the header of Image2, writing out to
Image2repaired.

@@

\subsubsection{Affine and Rigid Registration}
\begin{figure}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSanatomy.pdf} 
\itkcaption[ANTS Introduction]{The anatomy of an ANTS optimization and application of the resulting warping.}
\vspace{-0.1in}
\label{fig:aff}
\end{figure}
The most basic mapping between images is an affine mapping. The ANTS 
affine mapping syntax is shown in the affine mapping figure~\ref{fig:aff}. The
parameter breakdown of the ANTS call in the affine figure is:
\begin{itemize}
\item  ANTS -- the normalization program
\item 2 -- the expected image dimension
\item -m MI[...,...,1,32] -- the similarity metric "mutual information" with a 32-bin square joint histogram and weight of 1.  
Metric details will be discussed in a later section. 
\item -i 0 -- no deformable iterations (affine only).
\item -o ab -- the output prefix.  A specific extension may also be specified, e.g.   -o ab.mhd  .
\end{itemize}
The parameter breakdown of the WarpImageMultiTransform call in the affine figure is:
\begin{itemize}
\item WarpImageMultiTransform -- to apply the mapping output from ANTS to an image.
\item 2 -- the expected image dimension.
\item the image to be deformed is input next.
\item the deformed image output file name is next.
\item the transform itself is next -- if is affine and preceded by " -i ", we apply the inverse affine map.
\item the " -R " option dictates the domain you want to warp into -- usually the ``fixed'' image, unless using the inverse mapping, in which case one switches the role of fixed and moving images. 
\end{itemize}
Note that -i 0 (in the ANTS call) means that no deformable mapping is performed. The call "ANTS 2" means to treat the images as 2-dimensional. "ANTS 3" would be called for 3D images. The " -m " term defines the similarity metric. 
Note that no more than one image interpolation is used during optimization.  
That is, ANTS always references the original image before applying a transformation or series of transformations.

@@

\noindent{\bf Rigid Registration: }  ANTS will also perform rigid registration.  
To force the linear component of registration to be purely rigid (no scaling or shearing), 
then pass the parameter  \verb"--rigid-affine true " to 
the ANTS command line.   For example,  
\begin{verbatim}
ANTS 3 -m MI[AtlasHead.nii,crAnatomical.nii,1,32] 
-o TEST -i 0 --rigid-affine true 
\end{verbatim}
The jacobian of the resulting affine mapping should be unity.

\subsubsection{Deformable Registration}
Affine mapping is adequate when the
difference between images involves only rotation, scaling or shearing.
Other data requires more deformable mappings to capture shape
differences and find a good alignment of image features.
Figure~\ref{fig:antshier} shows how deformable mapping may improve the
correspondence of the deformed beetle to the ford image.
\begin{figure}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSElastToDiff.pdf} 
\itkcaption[ANTS Transformation Examples]{This example shows the degree to which the beetle (b.img) may be deformed to the ford (a.img) under different transformation models.  Left to right increases the degrees of freedom in the mapping and thus the registration accuracy.}
\vspace{-0.1in}
\label{fig:antshier}
\end{figure}
Most ANTS-based applications use symmetric diffeomorphic normalization.  However, 
ANTS also enables a simpler parameterization of a deformable mapping via a regularized 
vector space.  We term these types of transformations as ``elastic''.  The original Demons algorithm 
provides a classic example of using a regularized vector space for nonlinear registration.  
Caveats are that a regularized vector space may not preserve the underlying topology 
and may also prove too inflexible to capture the shape changes one is after.  Both of these 
shortcomings motivate the use of diffeomorphisms.

\noindent{\bf Elastic/Vector Space Transformations.} If we assume no affine transformation, then 
an elastic transformation involves computing a mapping from image $I(\x)$ to image $J(\x)$ 
through a deformation field $\disp( \x )$.  
\begin{figure}
\center \includegraphics[width=0.75\textwidth]{./Figures/PullBack.pdf} 
\itkcaption[ANTS Deformation]{A deformation in a digital domain.}
\label{fig:pull}
\end{figure}  
The deformation is defined in the physical space of the image 
and dictates the positional difference between corresponding features in the two images.  
Thus, if a feature defined at $I(\x)$ matches a feature in $J$ at position $\y$ then the deformation 
field at $\x$ should give $\disp ( \x )=\y-\x$.   
Such a deformation field may be applied to deform 
image $J$ into image $I$ by composing the mapping $J_{deformed}(\x)=J(\x+\disp(\x))$.  
In a perfect world, then $I(\x)=J_{deformed}(\x)$, though this is rarely the case.  Figure~\ref{fig:pull} 
visualizes the deformation of an image under this standard model. 
Gradient descent optimization of an elastic mapping may be summarized (crudely) as: 
\begin{align}
\text{Compute the similarity gradient:   }  \nabla E = \partial_{\disp} \Pi(I,J(\x+\disp(\x))). \nonumber \\
\text{Update the deformation field:  }  \disp(\x) \leftarrow \disp(\x)+\delta \nabla E. \nonumber \\ 
\text{Regularize the deformation field:  }  \disp(\x) \leftarrow G_\sigma \star \disp(\x),
\end{align}
where $\Pi$ is the similarity, $\delta$ is a gradient step length and $G_\sigma$ is a gaussian smoother.
This optimization is captured in the following ANTS command:
\begin{verbatim}
ANTS 3 -m PR[AtlasHead.nii,crAnatomical.nii,1,4] -r Gauss[0,3]  
-o ElasticTest.nii.gz -i 30x20x10  -t Elast[1.5] 
\end{verbatim}
Here, we use the PR metric (a cross-correlation implementation) with
window radius 4, weight 1 and gradient step length 1.5.  The
optimization will be performed over three resolutions with a maximum
of 30 iterations at the coarsest level, 20 at the next coarsest and 10
at the full resolution.  We use a Gaussian regularizer with a sigma of
3 that operates only on the deformation field and not on the similarity gradient, as 
0 is passed as the first parameter to \verb Gauss .  One may see the
correspondence, yet again, between the ANTS call and the optimization
scheme.  The optimization will stop when either the energy cannot be 
smoothly minimized or the maximum number of iterations is reached. 
BSpline regularization is also available in ANTS -- see the DMFFD 
section below.

\noindent{\bf Warping and Invertibility.}  ANTS uses physical space to
define mappings. The origin etc is in the coordinates of the world in
which the picture (e.g. MRI) was taken. Then, warping between physical
coordinates is relatively easy. Differences in bounding boxes etc
present no problem -- assuming you avoid inconsistent headers
i.e. origins, directions, data orientation. One may use PrintHeader to
check the data and run simple, fast tests (few iterations) to perform
sanity checks before running through loads of data.
An ANTS deformation consists of
a standard naming prefix plus a standard naming extension. We usually 
assume nii but other file types may be used. The standard naming
extension is Warpxvec.nii for the x component of the deformation with
similar naming for y and z. Each component is stored in a separate
scalar image. The value of a voxel of a deformation/warp component
image is the physical space displacement from that voxel. Note that
the inverse mapping is stored in InverseWarpxvec.nii and provides the
mapping in the opposite direction.  Note that an inverse -- in the 
digital domain -- is only approximate as shown in figure~\ref{fig:inv}.
\begin{figure}
\center \includegraphics[width=0.75\textwidth]{Figures/DeformationsInDigitalDomain.pdf} 
\itkcaption[Digital Invertibility]{Digital invertibility presents some limitations. 
Here, we see that invertibility is not exact but is gained only by interpolation.  
Thus, in three-dimensional scenarios in particular, there are fundamental limits to the 
degree of invertibility that may be achieved.  The second and third voxels -- from left -- 
in image A undergo an expansion by a factor of 2.  That is, under the mapping, 
2 voxels are mapped to 4.  This gives the definition of the Jacobian -- computed 
by  ANTSJacobian -- which is a unitless measure defined by the 
ratio of volumes.  Thus, $\mathcal{J}(\x)=V(\phi(\x))/V(\x)$ where $V$ represents 
the volume operation and $\x$, here, may be a small object.  Thus, if $\phi$ -- the mapping -- 
causes expansion, then $\mathcal{J}(\x) > 1$. }
\vspace{-0.1in}
\label{fig:inv}
\end{figure}  
A few important notes follow. 
(1) Deformation directionality: Warps/deformations applied to images occur in the opposite direction from warps/deformations applied to points. That is, if we map image B to A using 
\begin{verbatim}
ANTS 2 -m PR[A,B,1,2] -o OUTPUT 
\end{verbatim}
 then we get a deformation called \verb OUTPUTWarp+extensions ~and an affine transform called \verb OUTPUTAffine.txt . 
This composite mapping - when applied to B - will transform B into the space of A. 
However, if we have points defined in B that we want to map to A, 
we have to use \verb OUTPUTInverseWarp ~and the inverse of \verb OUTPUTAffine.txt .  
This is because the domain and range of the map's definition need to be dense in a way that is appropriate for the data type. 
Figure~\ref{fig:inv} illustrates the concept. 
(2) Older Image Formats: older image formats (e.g. Analyze) may not
have proper origin/offset/direction support! In these cases, we
recommend converting to nii and verifying that data overlays properly.
(3) Transorm Composition: Composition of transforms may be achieved with
ComposeMultiTransform.  (4) Warping with inverses and concatenations --
viable when using diffeomorphisms -- are described in
figure~\ref{fig:warp}.
 \begin{figure}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax2.pdf} 
\itkcaption[ANTS Warping]{Here, we detail how one applies 
a deformation and associated inverses.  We also see how 
WarpImageMultiTransform can be used to concatenate a series 
of transformations.}
\vspace{-0.1in}
\label{fig:warp}
\end{figure}

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\label{fig:antssyn}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization \cite{Avants2008b,Kim2008} -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

@@

\subsection{Diffusion Tensor Normalization}
The ImageMath program -- via TensorFA and TensorMeanDiffusion -- 
may derive scalar images from DTI.  Such images may be used to 
distortion correct DTI to T1 or to map DTI together.   See the 
ANTS/Scripts program called \verb antswithdt.sh ~for a 
ready to go example.   Figure~\ref{fig:dterr} shows what might 
happen if your tensor entries are stored in the wrong order. 
ANTS expects the nifti standard ordering of the DT six vector.
\begin{wrapfigure}{r}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{Figures/dtiordererr.pdf} 
\itkcaption[DT Entry Error]{The FA and mean diffusion of the tensor may be corrupted, 
as at right, if the order of the DT entries is wrong.}
\vspace{-0.1in}
\label{fig:dterr}
\end{wrapfigure}  
\subsection{Multivariate Normalization with ANTS}
\begin{figure}
\includegraphics[width=0.9\textwidth]{Figures/multivariatenorm.pdf} 
\itkcaption[Multivariate Normalization]{The current version of ANTS 
may perform normalization of multiple modalities by combining {\em intrasubject, intermodality} mappings with {\em intersubject, intramodality} maps to a 
group template.  In brain imaging, the intersubject maps are usually guided by the T1 component.}\vspace{-0.1in}
\label{fig:mvnorm}
\end{figure}
Multivariate normalization may be performed in two ways with ANTS.  
First, as shown in figure~\ref{fig:mvnorm}, intra-subject mappings may 
be used to conform all modalities into a common subject space defined by 
the ``anchor'' modality.  In brain imaging, this is usually the T1 component.

@@

The steps are: 
\itkcaption[ANTS Optimal Template]{This example may be recreated by the user.  
The shared anatomy -- 

@@

@@

\subsection{User/Label-Guided Normalization for Hippocampus Mapping}
See the link for details on hippocampus labeling:
\href{http://picsl.upenn.edu/ANTS/hipptutorial.php}{http://picsl.upenn.edu/ANTS/hipptutorial.php}.
\noindent{\bf Expectation-Based Point-Set Registration.}
Here, we apply the expectation-based point set registration method 
for mapping labeled points sets, as described in \cite{Pluta2008}.   ITK-SNAP may be used to label 
images and exported segmentation images may be input to the 
PSE metric below, as labeled data.  The Frown and Smile data is used 
as example.  This data is available in the 
ANTS/Examples/Data/ directory.
\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/frowntosmile.png}
\includegraphics[width=0.45\textwidth]{Figures/smiletofrown.png}
\itkcaption[Frown To Smile]{Expected output for the frown to smile 
shows a smooth, though large deformation.  The grids are overlaid on 
the deformed images. }
\label{fig:frown}
\end{figure}
\begin{verbatim}
ANTS 2 -o PSEtest  -i 91x70x55x40x30  -r Gauss[3,0.] -t SyN[0.2]
   -m  PSE[Frown.nii,Smile.nii,Frown.nii,Smile.nii,0.75,0.1,11,0,10] 
   -m  MSQ[Frown.nii,Smile.nii,1,0]    --number-of-affine-iterations 0

@@

The ANTS toolkit handles this situation by a constrained cost-function
masking approach. In short, the mapping within an excluded region (as
determined by an inclusive mask) is determined by the solution outside
of the region. To achieve this with ANTS, one would use a call similar
to this:
\begin{verbatim}
ANTS 3 -m PR[tp22_s1.nii,template.nii.gz,1,4] -i 50x20x0
   -o tp22map -t SyN[0.5] -x mask.nii.gz -r Gauss[3,0] 
\end{verbatim}
The additional parameter needed for this approach 
is the \verb -x ~ \verb mask.nii.gz ~ which 
is the mask defined in the "fixed" (or individual) image space.  
Mask values above 0.5 are considered non-outlier or non-lesioned regions.
The labeling may be performed with ITK-SNAP. 
Figure~\ref{fig:lesion} illustrates the approach.
\begin{figure}
\includegraphics[width=1\textwidth]{Figures/lesionstudy.pdf}
\itkcaption[ANTS Lesions]{Here, we see a semi-automated approach for lesion analysis
based on diffeomorphic normalization.  The user outlines the lesion
site in the subject space, as shown in (a).  Diffeomorphic mapping is
then used to deform the subject image in (a) into the space of the
reference template.  The deformed subject is shown in panel (b).
Panel (c) shows the reference template space with the estimated
position of the subject's lesion overlaid on the healthy tissue of the
template.  This approach enables one to compare the subject image
against prior information stored in the template space such as
anatomical labels or statistics on the appearance of lesioned and/or
normal tissue.}
\label{fig:lesion}
\end{figure}
Once the ANTS mapping is solved, 
one is able to estimate the location of the lesion in the template space by mapping the inverse lesion mask to the template.
To take the negative image of the inclusive mask 
and warp the mask to the template :
\begin{verbatim}
ImageMath 3 lesionmask.nii.gz Neg inclusivemask.nii.gz
WarpImageMultiTransform 3 lesionmask.nii.gz lesiontotemplate.nii.gz 
   -R template.nii.gz -i tp22mapAffine.txt tp22mapInverseWarp.nii
\end{verbatim}
More details on this approach are being prepared for publication.
See also: \verb http://www.ncrrn.org/papers/methodology_papers/sp_norm_kim.pdf .
